---
title: "Lecture2_racial_discrimination"
author: "Visseho Adjiwanou, PhD."
date: "2/17/2018"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Racial discrimination

Does racial discrimination exist in the labor market? Or should racial disparities in the unemployment rate be attributed to other factors such as racial gaps in educational attainment ?


## Slide with Bullets

- In response to newspaper ads, the researchers sent out r??sum??s of fictitious job candidates to potential employers
- Varied only the name of job applicant
- Other information are unchanged


## Slide with R Output

```{r racial discrimination}

## Set the work directory
getwd()


## Install and load the package "qss" that contains all data files from the book QSS - Imai Kosuke
# devtools::install_github("kosukeimai/qss-package")      # See the way the package is installed
# install.packages("rio")
library("qss")
library("tidyverse")
library("readr")          # load csv files
library("haven")          # for stata, spss and sas file
library("rio")

data("resume", package = "qss")
# resume <- read_csv("resume.csv")        # What you will do to import the data (remember your directory)
resume

```


## Quick observation of tha data

```{r racial discrimination}

## Subsetting the data
head(colnames(resume))      # show the variables names
head(resume)                # show first several observations
summary(resume)
glimpse(resume)

table(resume$race, resume$call)
table(resume$sex, resume$race)

# Another way       - By using the tidyr
race_call_tab <- 
  resume %>%
  group_by(race, call) %>%
  count()
race_call_tab

# All people who were female
people_female <- resume[resume$sex < 2, ]
summary(people_female)

# Complete case
no_missing_data <- resume[complete.cases(resume), ]

# subsample

sex_call <- resume[, c("sex", "call")]
sex_call

# number of column
ncol(sex_call)
ncol(resume)

```


## Proportion call back by race


```{r, racial discrimination}

# Contengency table - cross tabulation 

table(resume$race, resume$call)
table(resume$sex, resume$call)

race.call.tab <- table(race = resume$race, call = resume$call)    # if you want to re-use the table later
race.call.tab

addmargins(race.call.tab)

# overall call back rate: total callbacks divided by the sample size

sum(race.call.tab[, 2]) / nrow(resume)        # To what refers to race.call.table[, 2]
race.call.tab[1,2]

# Call back race for each race

(black.call <- race.call.tab[1, 2] / sum(race.call.tab[1,]))    # Black

(white.call <- race.call.tab[2, 2] / sum(race.call.tab[2, ]))   # White

white.call - black.call

# Overall call back simply

mean(resume$call)         # Because call is a dummy variable

```


## Subsetting the data to produce the same results

```{r, racial discrimination}

## What is the call back for each race when one want to use mean

resume$race

# Callback rate for black-sounding names
mean(resume$call[resume$race == "black"])

# Callback rate for white-sounding names
mean(resume$call[resume$race == "white"])


``` 

## Subsetting the data 

Cleaning/Reshapring/???Wrangling??? data is a core task of quantitative social scientist more broadly. A recent New York Times Article suggests 80% of data scientists' time is spent cleaning data, while only 20% of their time is spent analyzing it.



## Hawthorne effect

The howtorn effect refers to the phenomenon where study subjects behave differently because they know they are being observed by researchers.

```{r, hawthorne}
social <- data("social", package = "qss")
social

head(social)


```




