---
title: "Section5_TP5_Pression sociale"
author: "Visseho Adjiwanou, PhD."
date: "10/3/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Social pressure and the Hawthorne effect

Does social pressure within neighborhood increases participation (especially during a primary election)?

The Hawthorne effect refers to the phenomenon where study subjects behave differently because they know they are being observed by researchers.

## load the library and the data

```{r social}

## Set the work directory
getwd()
#setwd()

## Install and load the package "qss" that contains all data files from the book QSS - Imai Kosuke
# devtools::install_github("kosukeimai/qss-package")      # See the way the package is installed
# install.packages("rio")
library("qss")
library("tidyverse")
library("readr")          # load csv files
library("haven")          # for stata, spss and sas file
library("rio")


data("social", package = "qss")
write.csv(social, file = "social.csv")

social
head(social)
summary(social)

```


## Turnout and SATE

```{r, turnout}
# turnout of each group

message.tab <- table(social$messages, social$primary2006)
addmargins(message.tab)

# rate participation
sum(message.tab[, 2]) / nrow(social)
sum(message.tab[, 2]) / sum(message.tab[,])     # Another way to do it
mean(social$primary2006)

# participation civil duty
(civil.duty <- message.tab[1, 2] / sum(message.tab[1, ]))
civil.duty

# participation control group
control <- message.tab[2, 2] / sum(message.tab[2, ])
control

# participation hawthorne group
hawthorne <- message.tab[3, 2] / sum(message.tab[3, ])
hawthorne

# participation neighbors group
neighbor <- message.tab[4, 2] / sum(message.tab[4, ])
neighbor


# SATE for each treatment group

civil.duty - control
neighbor - control
hawthorne - control

```


## Use of the function    (tapply)

What we have done is a very long process. we can do that with less steps by taking advantage of the function in R. here, we are going to use the function tapply.

```{r social }

# turnout of each group
tapply(social$primary2006, social$messages, mean)

# turnout of the control group

mean(social$primary2006[social$messages == "Control"])

# SATE

tapply(social$primary2006, social$messages, mean) - mean(social$primary2006[social$messages == "Control"])

```


## How can we be sure that the treatment groups are similar to the control group?


```{r, social}

# Are all four groups are similar before the treatment?

# Comparison of age 
social$age <- 2006 - social$yearofbirth

tapply(social$age, social$messages, mean)

# Comparison of previous election
tapply(social$primary2004, social$messages, mean)

# Comparison of hhsize
tapply(social$hhsize, social$messages, mean)


# comparison of sex

tapply(social$sex, social$messages, mean)

social$sex[social$sex == "Female"] <- "0"
social$sex[social$sex == "Male"] <- "1"

tapply(social$sex, social$messages, mean)

```




